Non-word repetition in Yélî Dnye

### 2020-10-05 manuscript.tex
version that knits for sure

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfauthor={Alejandrina Cristia~\& Marisa Casillas},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Non-word repetition in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}~\& Marisa Casillas\textsuperscript{2}}
\date{}

\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d'Ulm, 75005 Paris, France. E-mail: \href{mailto:alecristia@gmail.com}{\nolinkurl{alecristia@gmail.com}}}

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. NWR scores are thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated without changes less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  draft discussion
\end{itemize}

\begin{itemize}
\tightlist
\item
  add information about proportion of mispronunciations that are real based on other papers
\item
  add effect sizes to all analyses
\item
  ? correct treatment of words whose mispronunciation is an English word:
  mispronunciation meaning
  361 two two
  362 two two
  879 no-pinêt English: no peanut
  908 nodedi English: no daddy
  933 no--dadi English: no daddy
  949 petrumi English: ``true''
  950 petrumi English: ``true''
  951 petrumi English: ``true''
  1373 wee English: where
  1374 wee English: where
  1375 wee English: where
  1379 wee English: where
  1380 wee English: where
  1388 wee English: where
  1394 wee English: where
  1396 wee English: where
  1398 wee English: where
  1399 wee English: where
  1402 wat English: what
  1405 wat English: what
  1411 wee English: where
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and to assess the extent to which apparent phonological constraints found in the lexicon affect online production (Gallagher, 2014). It is also frequently used for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). However, it has been seldom used outside of Europe and North America (with exceptions including Gallagher, 2014, and Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020)).

@@

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

While all consonants can appear in the first syllable of the word, the second syllable onset consonant in bisyllabic roots is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see Levinson (n.d.) for details).

@@

The basic structure of NWR involves the participant hearing a non-word, like ``bilik'', which they need to repeat back without changing anything that is contrastive in the language. For instance, in English, a response of ``bilig'' or ``pilik'' will be scored as incorrect; a response ``biilik'', where the vowel is lengthened without change of quality would be scored as correct, because English does not have a pure length contrast.

@@

A first list of candidate items was generated in 2018 by selecting simple consonants (``p'', ``t'', ``d'', ``k'', ``m'', ``n'', ``w'', ``y'') and vowels (``i'', ``o'',``u'', ``a'',``e'') that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described (Levinson, n.d.), there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset \textbf{XXX }mc** please fill in** and an Olympus \textbf{XXX }mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}). This item is removed from analyses below.

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items `spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.

The child was donned a headset (\textbf{xx }mc** please fill in** for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring (e.g., Cristia et al., 2020). We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being ``bilik'', and imagine a response ``bilig''. This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20\% (20\% of phonemes were substituted or deleted).\\

@@

A total 55 children were tested, from 34 of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=8), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.96 years (range 3.90-11 years). Among these, there were 32 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and 2 for whom this information is missing. Maternal years of education averaged 8.24 years (range 6-12 years; 2 children had this information missing).\footnote{Precise years of education can be hard to count. This is particularly the case for schooling beyond grade 8, which takes place off-island, which can be due to an irregularity in ability to pay school fees, school opening times, familial obligations, and more. We asked parents to report their highest completed level of education and inferred the approximate number of years that entails under ideal conditions, which results in many esimates around typical ``completion'' times in between.} In terms of birth order, 5 were first borns, 4 second, 2 third, 7 forth, 5 fifth, 1 sixth, and NA did not have this information.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig1-first_vs_others-1.pdf}
\caption{(\#fig:Fig1-first\_vs\_others)NWR scores for individual participants averaging separately their first attempts and all other attempts.}
\end{figure}

We first checked whether NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, NWR scores for subsequent repetitions (M = 40, SD = 26)
were on average lower than first ones (M = 64, SD = 15), t(38) = 6.28, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 65\% (SD = 15\%). Scores based on phonemes are even higher M = 78\% (SD = 9\%). The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

In this analysis, we were interested in variation in NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.82, SE ß = 0.18, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.33, SE ß = 0.13, p = 0.01).

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.52, =0.00). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p = 0.00) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of word-level NWR scores, but corpus frequency (ß = 0, SE ß = 0.25, p = 1.00) was not.

\textbf{use proportion of correctly produced rather than number; DONE}

\textbf{a2m use first attempts rather than the whole data set: having second thoughts. To be clear: the argument for using first attempts only is that all of the other analyses are on kids' first attempts, rather than all of their attempts, so we should use just first attempts here because it may confuse readers \& it is more consistent. BUT there are fewer mistakes this way, which is Options 1 and 2 below has lots of zeros\ldots{} CF Option 3 further down}

\textbf{a2m probably turn the paragraph on substitutions into a table: having second thoughts, the table would look just like a para, no?}

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels, 30 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (110 oral vowels produced with a different quality; 13 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 115) than vice versa (2). Simple consonants were mispronounced as other simple consonants quite frequently (69 simple consonants mispronounced as other simple ones, compared to 33 complex ones).

@@

\textbf{(MC: but we can try and do a cursory analysis based on the corpora we have from Steve and my transcription of naturalistic interactions!)}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{NWR in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}\ \& Marisa Casillas\textsuperscript{2}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics}
\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Non-word repetition in Yélî Dnye}

\date{}

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. NWR scores are thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated without changes less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.
}

@@

\begin{itemize}
\tightlist
\item
  check again if all fields are explained
\item
  set up OSF
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\end{itemize}

\begin{itemize}
\tightlist
\item
  draft discussion
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and to assess the extent to which apparent phonological constraints found in the lexicon affect online production (Gallagher, 2014). It is also frequently used for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). However, it has been seldom used outside of Europe and North America (with exceptions including Gallagher, 2014, and @cristia2020infant), and mostly with languages having phonological inventories that are small to moderate in size (Maddieson, 2005).

@@

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., \enquote{ndap} (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., \enquote{ndapî} /\enquote{ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî} (phone) /\enquote{pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala} (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

@@

The basic structure of NWR involves the participant hearing a non-word, like \enquote{bilik}, which they need to repeat back without changing anything that is contrastive in the language. For instance, in English, a response of \enquote{bilig} or \enquote{pilik} will be scored as incorrect; a response \enquote{biilik}, where the vowel is lengthened without change of quality would be scored as correct, because English does not have a pure length contrast.

@@

A first list of candidate items was generated in 2018 by selecting simple consonants (\enquote{p}, \enquote{t}, \enquote{d}, \enquote{k}, \enquote{m}, \enquote{ń}, \enquote{w}, \enquote{y})
and vowels (\enquote{i}, \enquote{o},\enquote{u}, \enquote{a},\enquote{e}) that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described (Levinson, n.d.), there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a Shure SM10A XLR dynamic headband microphone and an Olympus WS-832 stereo audio recorder (using an XLR to mini-jack adapter) from the written form presented together with the same item orally by the second author. The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended \enquote{yî} /yXX/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from analyses below.

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items \enquote{spreading} through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.

The child was donned a headset (a Shure SM10A or WH20 XLR with a dynamic microphone on a headband, most children using the former), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring (e.g., Cristia et al., 2020). We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being \enquote{bilik}, and imagine a response \enquote{bilig}. This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20\% (20\% of phonemes were substituted or deleted).\\

@@

A total 55 children were tested, from 38 of families, in four hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=8), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.98 years (range 3.92-11.03 years). Among these, there were 34 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and 0 for whom this information is missing. Maternal years of education averaged 8.22 years (range 6-12 years; 0 children had this information missing).\footnote{Precise years of education can be hard to count. This is particularly the case for schooling beyond grade 8, which takes place off-island, which can be due to an irregularity in ability to pay school fees, school opening times, familial obligations, and more. We asked parents to report their highest completed level of education and inferred the approximate number of years that entails under ideal conditions, which results in many esimates around typical \enquote{completion} times in between.} In terms of birth order, 6 were first borns, 5 second, 2 third, 7 forth, 5 fifth, 1 sixth, and NA did not have this information.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig1-first_vs_others-1.pdf}
\caption{(\#fig:Fig1-first\_vs\_others)Whole-item NWR scores for individual participants averaging separately their first attempts and all other attempts.}
\end{figure}

We first checked whether whole-item NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, whole-item NWR scores for subsequent repetitions (M = 40, SD = 26)
were on average lower than first ones (M = 64, SD = 15), t(38) = 6.28, p = 0; Cohen's d = 1.14. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 65\% (SD = 15\%), Cohen's d = 4.34. Scores based on phonemes are even higher M = 78\% (SD = 9\%), Cohen's d = 8.50. The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

In this analysis, we were interested in variation in whole-item NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 850 observations, from 40 children producing in any given trial one of 41 potential target words. The analysis revealed a main effect of age (ß = 0.33, SE ß = 0.13, p = 0.01); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = 0.82, SE ß = 0.18, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r r(31) = 0.63.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence is actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020) by counting the number of word types in which they occurred, and applied the natural logarithm.\footnote{We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.} As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequencies estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.52, =0.00). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = 0.47. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.83, SE ß = 0.27, p = 0.00) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of whole-item NWR scores, but corpus frequency (ß = 0.00, SE ß = 0.25, p = 1.00) was not.

\hypertarget{patterns-in-nwr-mispronunciations}{%
\subsubsection{Patterns in NWR mispronunciations}\label{patterns-in-nwr-mispronunciations}}

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels, 30 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (110 oral vowels produced with a different quality; 13 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mńlwyvdgptkfhch, 115) than vice versa (2). Simple consonants were mispronounced as other simple consonants quite frequently (68 simple consonants mispronounced as other simple ones, compared to 33 complex ones).

@@
